---
title: "STATS PROJECT 2022-Multi-stuff-metabolite"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
pagetitle: Habits
editor_options:
  chunk_output_type: console
---

```{r, massage=FALSE,echo=FALSE, include=FALSE}
library("lme4")
library("lattice")
library("minqa")
library("dplyr")
library("tidyverse")
library("janitor")
library("dplyr")
library("lcmm")
library("Matrix")
library("stringr")
library("reshape2")
library("ggpubr")
#library('ClusterR')
library('cluster')
library('caret')
library('factoextra') # clustering visualization
library('dendextend') # for comparing two dendrograms
library(dplyr) 
library(purrr)
library("modelr")
library(gtools)

```


```{r echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
setwd("~/Documents/repos/Dissertation_F/Dissertation-F")
raw_data = read.csv("Acute_Metabolites and Lipids.csv") # nolint
dim(raw_data)
data =raw_data[2:534]
length(unique(data$Study.ID))
glimpse(data)

View(data)
```


1- cleaning
```{r,include=FALSE}
#install.packages("janitor")
cleaned_data = clean_names(data)
cleaned_data$meal =as.numeric(as.factor(cleaned_data$meal))
cleaned_data$time_mins = as.numeric(cleaned_data$time_mins)
cleaned_data$gender = as.factor(cleaned_data$gender)
cleaned_data$patient_id=as.factor(cleaned_data$patient_id)
cleaned_data$meal = as.factor(cleaned_data$meal)
cleaned_data$ethnicity= as.factor(cleaned_data$ethnicity)

table(cleaned_data$ethnicity)
view(cleaned_data)

cleaned_data %>%  filter(study_id =="4203" )
```

```{r,include=FALSE}
test.data = cleaned_data %>%  select(!c(sequence,left_over,batch,no,visit_nr,initials,age_yrs,weight_kg,height_m,bmi_kg_m2,mri_pancreas_fat_mr_opsy_thresholded,participant))
glimpse(test.data)
length(test.data$patient_id)
unique(test.data$patient_id)
length(unique(test.data$study_id))


length(dim(test.data[7:521]))
```

*Taking only the metabolites with significant p-values*
The forloop for all metaboloutes -test loop only 5 metabolotes 
```{r, include=FALSE,warning=TRUE,results='hide'}
test.data$time_mins = as.factor(test.data$time_mins)
var_list <- names(test.data)[8:100]
length(var_list)
mod1.0 <- lapply(var_list, function(x) {
 lmer(substitute(i ~ time_mins+meal+(1|study_id), 
 list(i = as.name(x))), data = test.data, na.action=na.exclude)
})

mod2.0 <- lapply(var_list, function(x) {
 lmer(substitute(i ~ time_mins*meal+(1|study_id), 
 list(i = as.name(x))), data = test.data, na.action=na.exclude)
})
multi_anova = mapply(anova, mod1.0, mod2.0, SIMPLIFY=FALSE)
p_values = as.list(NULL)
p_values_like = as.list(NULL)
for(i in 1:length(multi_anova)){
   p_values[i]= multi_anova[[i]]$`Pr(>Chisq)`[2]
}

#p_values
p_values_like =p_values[p_values <= 0.05]  

pvals = which(p_values <0.05)
metabolite_sig_df = test.data[,pvals]
metabolite_sig_df = cbind( test.data[3:7],metabolite_sig_df)
# colnames(metabolite_sig_df)
# metabolite_sig_df$time_mins
metab_df = metabolite_sig_df[,4:8]

all_metab_sig = metabolite_sig_df %>%  select(!c(gender,ethnicity,patient_id,id_meal_time))

all_metab_sig$studyid_time = paste(all_metab_sig$study_id, "-", all_metab_sig$time_mins)

all_metab_sig = all_metab_sig %>%  select(studyid_time,study_id, meal, time_mins,diaminopimelic_acid : x4_hydroxyproline)
length(all_metab_sig$study_id)
length(all_metab_sig$study_id)

as.character(colnames(all_metab_sig[5:36]))

xx =all_metab_sig %>% group_by(study_id, time_mins)

review_tab1 = all_metab_sig %>% select( time_mins, meal, diaminopimelic_acid : x4_hydroxyproline)
head(review_tab1)



review_tab1$diaminopimelic_acid = as.numeric(review_tab1$diaminopimelic_acid)
review_tab1$time_mins = as.numeric(review_tab1$time_mins)
review_tab1$meal = as.numeric(review_tab1$meal)


str(review_tab1)
meal_one = review_tab1 %>%  filter( meal == "1")

v1 = aggregate(meal_one$time_mins~ meal_one$diaminopimelic_acid, FUN =  sum, data =  meal_one[1:3])
v1$`meal_one$time_mins` = as.factor(v1$`meal_one$time_mins`)

levels( v1$`meal_one$time_mins`) <- c("0", "30", "60", "120", "240")

v1 = aggregate(v1$`meal_one$diaminopimelic_acid` ~ v1$`meal_one$time_mins`, FUN = sum)
colnames(v1) = c("time", "diaminopimelic_acid")
v1 %>% pivot_wider(names_from = time, values_from = diaminopimelic_acid)

ggplot(v1, aes(time, diaminopimelic_acid, colour = time)) + 
  geom_point()

```




```{r}
rev = all_metab_sig %>% select(study_id, time_mins, meal, diaminopimelic_acid : x4_hydroxyproline)
colnames(rev)
summary(rev)
rev$time = as.factor(rev$time)
levels( rev$time) <- c("0", "30", "60", "120", "240")
ggplot(rev, aes(x=time_mins)) + 
  geom_line(aes(y = diaminopimelic_acid), color="steelblue", linetype="twodash") 

df <- rev %>%   select(time_mins, diaminopimelic_acid, creatine, uridine, acetylcholine, guanine,nitrate ) %>%
  gather(key = "variable", value = "value", -time_mins)
head(df)
cols = rainbow(12)

ggplot(df, aes(x =  time_mins, y = value)) + 
  geom_line(aes(color = variable, linetype = variable)) + 
   scale_color_viridis(discrete = TRUE, option = "D")+
  scale_fill_viridis(discrete = TRUE) +
  theme_minimal() +
  theme(legend.position = "bottom")+ labs(title = "Metabolites vs Time",
      x = "Time (Minutes)", y = "WPI values")

```




```{r}
 meal1 = rev%>%  filter( meal == "1")
meal1$diaminopimelic_acid = as.numeric(meal1$diaminopimelic_acid)
meal1$time_mins = as.numeric(meal1$time_mins)
meal1$meal = as.numeric(meal1$meal)
v11 = aggregate(meal1$time_mins, by = list(meal1$study_id, meal1$diaminopimelic_acid), FUN =  mean, data =  meal1[1:3])
colnames(v11) = c("study_id", "diaminopimelic_acid", "time")
v11$time = as.factor(v11$time)
levels( v11$time) <- c("0", "30", "60", "120", "240")


rev_mini = rev %>% select(c("study_id", "time_mins", "meal",'diaminopimelic_acid', 'creatine', "uridine", "acetylcholine", "guanine","ornithine" )) %>%  
  mutate_at(c("study_id", "time_mins", "meal",'diaminopimelic_acid', 'creatine', "uridine", "acetylcholine", "guanine","ornithine" ), as.numeric)
head(rev_mini)

rev$time_mins = as.numeric(rev$time_mins)
rev$meal = as.numeric(rev$meal)

###################33THIS IS THE UNIVARSAL DF############3
v12 = aggregate(rev_mini$time_mins, by = list(rev_mini$study_id,
                                              rev_mini$meal,rev_mini$diaminopimelic_acid,rev_mini$creatine, 
                                         rev_mini$uridine,rev_mini$acetylcholine,
                                         rev_mini$guanine, rev_mini$ornithine), FUN =  mean)
head(v12)
head(rev_mini)
colnames(v12) = c("study_id","meal",  "diaminopimelic_acid", "creatine", "uridine", "acetylcholine", "guanine", "ornithine", "time")
v12$time = as.factor(v12$time)
levels( v12$time) <- c("0", "30", "60", "120", "240")
head(v12)
# meatb1 = v12 %>% pivot_wider(names_from = meal, values_from = diaminopimelic_acid, values_fill = 0)


```



```{r}
 meal1_mini = rev_mini %>%  filter( meal == "1")

v_mini  = aggregate(rev_mini$time_mins, by = list(rev_mini$study_id, rev_mini$meal, rev_mini$diaminopimelic_acid, rev_mini$creatine, rev_mini$uridine, rev_mini$acetylcholine, rev_mini$guanine, rev_mini$ornithine), FUN =  mean, data =  rev_mini)

colnames(v_mini) = c("study_id","meal", "diaminopimelic_acid", "creatine", "uridine", "acetylcholine", "guanine", "ornithine", "time")
v_mini$time = as.factor(v_mini$time)
levels( v_mini$time) <- c("0", "30", "60", "120", "240")
head(v_mini)

v_mini = v_mini[, c(1,2, 9,  3,4, 6,7,8)]

meatb_all  = v_mini %>% pivot_wider(names_from = meal, values_from = c(diaminopimelic_acid,creatine,acetylcholine,  guanine, ornithine),values_fill = 0)

meatb_df = as.data.frame(meatb_all)
summary(metab_df)
test_data_long_tidyr <- pivot_longer(meatb_df[1:5], cols = starts_with("diaminopimelic_acid"))


# ggplot(data=test_data_long_tidyr,
#        aes(x=time, y=value, colour=name)) +
#   geom_line()
# 
# 
# 
# ggplot(data=test_data_long_tidyr,
#        aes(x=time, y=value, colour=name)) +
#    geom_line(color = "steelblue", size = 1) +
#   geom_point() + 
#     facet_wrap(study_id~.)


#aggregated so each time point has 1 value 
m1 = meatb_df[1:5]
sum_meatb1 =aggregate(cbind(m1$time, m1$study_id) ~ m1$diaminopimelic_acid_1+m1$diaminopimelic_acid_3+m1$diaminopimelic_acid_2 ,FUN = sum ,na.rm = TRUE)

colnames(sum_meatb1) = c("meal1", "meal3", "meal2", "time", "study_id")
sum_meatb1$time = as.factor(sum_meatb1$time)
levels(sum_meatb1$time ) <- c("0", "30", "60", "120", "240")
metab1_df  = sum_meatb1 %>%  filter(study_id == "4203")

```










```{r}

sum_meatb1$time = as.numeric(sum_meatb1$time)
sum_meatb1$meal1 = as.numeric(sum_meatb1$meal1)
aggregate(data = sum_meatb1[1:4], sum_meatb1$time ~ sum_meatb1$meal1,FUN = sum)
```


```{r}

v12$meal = as.factor(v12$meal)
v12$time = as.factor(v12$time)


ggplot(data = v12, aes(time,diaminopimelic_acid)) +
  geom_line(color = "steelblue", size = 1) +
  geom_point() + 
  labs(title = "",
       y = "Diaminopimelic_acid ", x = "Time (Min)") + 
  facet_wrap(.~meal)


v12_2 = v12 %>% filter(study_id =="4201" | study_id =="4203"  | study_id =="4204" |study_id =="4208"  )
ggplot(data= v12_2,
       aes(x=time, y= diaminopimelic_acid, colour= meal)) +
   geom_line(color = "steelblue", size = 1) +
  geom_point() + 
 labs(title = "",
      x = "Time (Minutes)", y = " Diaminopimelic Acid",  color = "Preload Beverages\n") +
  scale_color_manual(labels = c("Meal 1", "Meal 2", "Meal 3"), values = c("red", "green","blue")) +    facet_wrap(study_id~.)

```


```{r}

#THIS IS THE UNIVARSAL DF####
ggplot(data = v12, aes(time,creatine)) +
  geom_line(color = "steelblue", size = 1) +
  geom_point() + 
  labs(title = "",
       y = "Creatine ", x = "Time (Min)") + 
  facet_wrap(.~meal)



#INDIVIDUAL METAB INFO 
m2 = meatb_df[c(1,2,6,7,8)]
colnames(m2) = c("study_id", "time", "meal1", "meal3", "meal2")
metab2  <- pivot_longer(meatb_df[c(1,2,6,7,8)], cols = starts_with("creatine"))
#aggregated so each time point has 1 value 

sum_metab2 =aggregate(cbind(m2$time, m2$study_id) ~ m2$creatine_1+m2$creatine_2+m2$creatine_3,FUN = sum ,na.rm = TRUE)

colnames(sum_metab2) = c("meal1", "meal3", "meal2", "time", "study_id")
sum_metab2$time = as.factor(sum_metab2$time)
levels(sum_metab2$time ) <- c("0", "30", "60", "120", "240")
metab2_df  = sum_metab2 %>%  filter(study_id == "4203")

str(metab2)








metab2_2  = metab2 %>% filter(study_id =="4201" | study_id =="4203"  | study_id =="4204" |study_id =="4208"  )


ggplot(data= metab2_2,
       aes(x=time, y=value, colour=name)) +
   geom_line(color = "steelblue", size = 1) +
  geom_point() + 
 labs(title = "",
      x = "Time (Minutes)", y = " Creatine",  color = "Preload Beverages\n") +
  scale_color_manual(labels = c("Meal 1", "Meal 2", "Meal 3"), values = c("red", "green","blue")) +    facet_wrap(study_id~.)



```

2- Set up the data frame 
*Need to select number of metabolite you want & check the names* 
```{r}
metab_names= names(all_metab_sig[,5:36])
selected_metab = do.call(cbind,all_metab_sig[5:36])
colnames(selected_metab) 
glimpse(selected_metab)
#selected_metab = as.numeric(selected_metab)


glimpse(all_metab_sig)
all_metab_sig$time_mins =as.character(all_metab_sig$time_mins)

 
 
metab_df = do.call(cbind,list(all_metab_sig$study_id,all_metab_sig$time_mins,all_metab_sig$meal,selected_metab))
colnames(metab_df)[1]  <- "study_id"
colnames(metab_df)[2]  <- "time_mins"
colnames(metab_df)[3]  <- "meal"
glimpse(metab_df)

# colnames(metab_df)[colnames(metab_df)=="all_metab_sig$study_id"] <- "study_id"
# colnames(metab_df)[colnames(metab_df)=="all_metab_sig$time_mins"] <- "time_mins"
# colnames(metab_df)[colnames(metab_df)=="all_metab_sig$meal"] <- "meal"

metab_df =  as.data.frame(metab_df)
head(metab_df)
```

*combine to make reshaping easier for whole dataframe*
```{r}
#COmbine only needed coloumns 
metab_df$studyid_time = paste(as.character(metab_df$study_id), "-", as.character(metab_df$time_mins))

#remove unneeded values cuz you combined them in new coloumn 
metab_df = metab_df %>%  select(!c(time_mins,study_id))
head(metab_df)
metab_df$meal = as.factor(metab_df$meal)
metab_df$diaminopimelic_acid = as.double(metab_df$diaminopimelic_acid)
metab_df$alpha_l_fucose = as.double(metab_df$alpha_l_fucose)
```

```{r}
each_metab = data.frame(matrix(NA, ncol = 5,nrow = 358))
cols1 <- c('meal', 'studyid_time')
cols2 <- setdiff(colnames(metab_df), cols1)
```


 make sure all variables are in the same class
```{r}
each_metab =lapply(cols2, function(x) metab_df[, c(cols1, x)])
each_metab =lapply(cols2, function(x) metab_df[, c(cols1, x)])
glimpse(each_metab)
class(colnames(selected_metab))
names_em =lapply(each_metab, function(x) names(x[3]))
names_em = as.character(names_em)
```
Checkj order and same 
```{r}
test_match_order <- function(x,y) {

  if (all(x==y)) print('Perfect match in same order')

  if (!all(x==y) && all(sort(x)==sort(y))) print('Perfect match in wrong order')

  if (!all(x==y) && !all(sort(x)==sort(y))) print('No match')
}
s = colnames(selected_metab)
b= names_em
test_match_order(s,b)
```


```{r}
f =function(x) as.numeric(x[,3])
f(each_metab[[1]])
lapply(each_metab, function(x) as.numeric(x[,3]))  #how to 
```

```{r}
#if the $ glutamylthreonine: chr [1:358]  is a charchracter need to change it to numeric 
# class(each_metab[[6]]$acetylcholine)
each_metab[[3]]$uridine = as.numeric(each_metab[[3]]$uridine)
each_metab[[4]]$glutamylthreonine = as.numeric(each_metab[[4]]$glutamylthreonine)
each_metab[[5]]$x1_methylpyrrolinium = as.numeric(each_metab[[5]]$x1_methylpyrrolinium)
each_metab[[6]]$acetylcholine = as.numeric(each_metab[[6]]$acetylcholine)
each_metab[[7]]$l_glutamine = as.numeric(each_metab[[7]]$l_glutamine)
each_metab[[8]]$n_6_methyllysine = as.numeric(each_metab[[8]]$n_6_methyllysine)
each_metab[[9]]$suberic_acid = as.numeric(each_metab[[9]]$suberic_acid)
each_metab[[10]]$guanine = as.numeric(each_metab[[10]]$guanine)
each_metab[[11]]$myo_inositol = as.numeric(each_metab[[11]]$myo_inositol)
each_metab[[12]]$l_beta_aspartyl_l_glutamic_acid = as.numeric(each_metab[[12]]$l_beta_aspartyl_l_glutamic_acid)
each_metab[[13]]$aspartyl_hydroxyproline = as.numeric(each_metab[[13]]$aspartyl_hydroxyproline)
each_metab[[14]]$x_s_2_hydroxybutyric_acid = as.numeric(each_metab[[14]]$x_s_2_hydroxybutyric_acid)
each_metab[[15]]$x2_hydroxyadipic_acid = as.numeric(each_metab[[15]]$x2_hydroxyadipic_acid)
each_metab[[16]]$x3_hexenedioic_acid = as.numeric(each_metab[[16]]$x3_hexenedioic_acid)


each_metab[[17]]$l_theanine = as.numeric(each_metab[[17]]$l_theanine)

each_metab[[18]]$oxoglutaric_acid= as.numeric(each_metab[[18]]$oxoglutaric_acid)

each_metab[[19]]$aspartyl_l_proline = as.numeric(each_metab[[19]]$aspartyl_l_proline)

each_metab[[20]]$pyroglutamic_acid = as.numeric(each_metab[[20]]$pyroglutamic_acid)

each_metab[[21]]$l_aspartic_acid = as.numeric(each_metab[[21]]$l_aspartic_acid)

each_metab[[22]]$creatine = as.numeric(each_metab[[22]]$creatine)

each_metab[[23]]$l_glutamic_acid = as.numeric(each_metab[[23]]$l_glutamic_acid)
each_metab[[24]]$l_lysine = as.numeric(each_metab[[24]]$l_lysine)
each_metab[[25]]$x2_aminomuconic_acid_semialdehyde = as.numeric(each_metab[[25]]$x2_aminomuconic_acid_semialdehyde)

each_metab[[26]]$methionine_sulfoxide = as.numeric(each_metab[[26]]$methionine_sulfoxide)

each_metab[[27]]$x2_keto_6_acetamidocaproate = as.numeric(each_metab[[27]]$x2_keto_6_acetamidocaproate)
each_metab[[28]]$thiocyanate= as.numeric(each_metab[[28]]$thiocyanate)
each_metab[[29]]$nitrate = as.numeric(each_metab[[29]]$nitrate)

each_metab[[30]]$x1_2_3_trihydroxybenzene = as.numeric(each_metab[[30]]$x1_2_3_trihydroxybenzene)

each_metab[[31]]$ornithine = as.numeric(each_metab[[31]]$ornithine)

each_metab[[32]]$x4_hydroxyproline = as.numeric(each_metab[[32]]$x4_hydroxyproline)





```

make sure there are no coloumns that have all NA
```{r}
each_metab_dfdc= lapply(each_metab, function(x) {melt(x)})
glimpse(each_metab_dfdc)

each_metab_dfdc=lapply(each_metab_dfdc, function(x) dcast(x,  meal ~ studyid_time, mean))
length(each_metab_dfdc[[18]])
```


```{r}
x = each_metab_dfdc[[18]][1]
y = each_metab_dfdc[[18]][2:121]
ggplot(each_metab_dfdc[[18]], aes(x = meal)) + geom_point()
```




3- Final_DF
```{r}
final_df =  data.frame(matrix(NA, ncol = (length(each_metab_dfdc)*3+2), nrow = 96))     # Creating data containing NA
rep_names =rep(colnames(selected_metab))
rep_vals = as.numeric(rep(1:3))


 last_part = character() 
first_part = character()
for(k in rep_names){
 for(n in rep_vals){
          output1 = paste("_", k, sep = "")
             last_part = c(last_part, output1)

              output2  = paste("meal", sep = "", sprintf("%02d",n))
                first_part = c(first_part, output2)
}

}

last_part
first_part
all_names = mapply(paste, first_part, last_part)

all_names_vec = as.vector(all_names) 

colnames(final_df) = c("study_id", "time_mins", all_names_vec)

for(i in 1:length(each_metab_dfdc)){
  #gather the values with 0 timepoints
   zero_timepoint <- grep("- 0$", colnames(each_metab_dfdc[[i]]),value = TRUE)
   zero_timepoint  <- each_metab_dfdc[[i]][,zero_timepoint]

   #gather the values for other times 
    nonzero_timepoints <- grep("- 30$|- 60$|- 120$|- 240$", colnames(each_metab_dfdc[[i]]),value = TRUE)
    nonzero_timepoints= each_metab_dfdc[[i]][,nonzero_timepoints]
    
    #combine timepoints
    all_timepoints_ids <- match(substr(colnames(nonzero_timepoints), 0, 4), substr(colnames(zero_timepoint), 0, 4)) #[-1]
    zero_timepoint2 <- zero_timepoint[, all_timepoints_ids]
    df_diff_time = nonzero_timepoints - zero_timepoint2
    df_diff_time$meal = each_metab_dfdc[[i]]$meal

    
    #reshape df- the "metabolite" is just a place holder for actual name 
    final_df_all = df_diff_time %>% pivot_longer(!meal, names_to = "studyid_time", values_to = "metabolite")
    final_df_all=melt(final_df_all)
    final_df_all =dcast(final_df_all, studyid_time ~meal, mean)
    
    #clean up the names and remove the two NAN
    # colnames(final_df_all)[colnames(final_df_all)=="1"] <- "meal01"
    # colnames(final_df_all)[colnames(final_df_all)=="2"] <- "meal02"
    # colnames(final_df_all)[colnames(final_df_all)=="3"] <- "meal03"
    final_df_all = final_df_all%>%  separate(studyid_time, c("study_id", "time_mins"))

    #store answer 
    # print(final_df_all[i])
    final_df[,1:2] = final_df_all[,1:2]
    # final_df[,i*3:(i*3+2)] = final_df_all %>% select(meal01,meal02,meal03)

    #print((i*3):(i*3+2))
    final_df[,(i*3):(i*3+2)] = final_df_all[,3:5]

}

glimpse(final_df)
any(duplicated(names(final_df))) #tested if colnames are unique FALSE --is wat u want

# mapply(function(X,Y) {
#   sapply(1:10, function(row) cor(X[row,], Y[row,]))
#   }, X=listA, Y=listB)

# mapply(function(x,y){paste("meal", sprintf("%02d", var_num), min_names)}, x = var_num, y = min_names)

final_df

```

Reorder the time into 30-60-120-240
```{r}
 final_df$study_id = as.integer(final_df$study_id)
final_df$time_mins = as.character(final_df$time_mins)
final_df$time_mins = as.numeric(final_df$time_mins)
# final_df$meal03_uridine = as.integer(final_df$meal03_uridine)
glimpse(final_df)

final_df=final_df %>%  mutate(final_df, time_mins_st = factor(time_mins, levels = c("30", "60", "120", "240")))   %>%  group_by(study_id) 
final_df =final_df %>%  group_by(study_id) %>%  arrange(study_id,time_mins_st)
```


DO THIS PART THE LAST: omit and create avg col 
```{r}
final_df= na.omit(final_df)
final_df =final_df %>%  group_by(study_id) %>%  mutate(avg_time = mean(time_mins))
glimpse(final_df)

```

```{r}
final_df <- as.data.frame(lapply(final_df, unlist))


final_df =as.data.frame(final_df)
class(final_df)
glimpse(final_df)
final_df_copy = final_df
```

```{r}
mini_df_list = list()
final_df4 = as.data.frame(final_df[3:34])
colnames(final_df4)
positions <- c(22:24)
df_plot = final_df4 %>% select(positions)
colnames(df_plot)
new_minidf = cbind(final_df[1:2], df_plot) 
colnames(new_minidf)
lt <- c("meal02" = "dash", "meal01" = "dotted","meal03" = "solid" )
p = ggplot(new_minidf, aes(study_id)) +     
  geom_line(aes(y = colnames(new_minidf[4]), linetype = "meal01",color = "red", group=1, )) + #mela 02 
 geom_line(aes(y = colnames(new_minidf[5]), linetype = "meal03",color = "green", group=1)) +  #meal 03 
  geom_line(aes(y = colnames(new_minidf[3]), linetype = "meal02", color = "blue",group=1)) #meal 01
    


```



# Mini DF
```{r}
final_df3 <- final_df %>% dplyr:: select(time_mins, study_id,starts_with("meal03"))
glimpse(final_df3)
dim(final_df3)
final_df3[1:2]
colnames(final_df3[3:34])
mini_df_list = list()
final_df4 = as.data.frame(final_df3[3:34])
colnames(final_df4[1:3])


```


*Best Link function*
```{r,cache=TRUE}
positions <- seq(1:32)
final_df5 =final_df4 %>% select(positions)
colnames(final_df5)
new_minidf = cbind(final_df3[1:2], final_df5) 
link_fun = vector(mode="list", length=length(positions))

# get the AIC  for all link functions 
class_lat = c(2,3)
link_vec = c("splines","beta")
for(i in 1:length(positions)){
    link_fun[[i]]<-rep(NA,2)
  for(k in 1:length(link_vec)){
    names(new_minidf)[3] <- "metab"
    new_minidf[,3] = final_df5[i]
    init = lcmm(metab ~time_mins, subject = "study_id", link = link_vec[k],random  = ~time_mins,
                data = new_minidf)
     link_fun[[i]][k]  <-  init$AIC
  }
}

link_fun

# Get the best latent class for initial model 
min_aic2 = list()
best_link =  data.frame(matrix(NA, ncol = length(link_fun)))
colnames(best_link) = paste("M-", 1:length(positions), sep = "")
for(i in 1:length(link_fun)){
    names(link_fun[[i]]) <- link_vec
    min_aic2[[i]] = which.min(link_fun[[i]])
    best_link[,i] =  link_vec[as.numeric(min_aic2[[i]])]
}
best_link



#Get the best Number of latent class *

output_aic_f = vector(mode="list", length=length(positions))

class_lat = c(2,3)
# link_vec = c("splines","beta")
for(i in 1:length(positions)){
    link_fun[[i]]<-rep(NA,2)
     for(c in 1:length(class_lat)){
     names(new_minidf)[3] <- "metab"
     new_minidf[,3] = final_df5[i]
     init = lcmm(metab ~time_mins, subject = "study_id", link = as.character(best_link[i]),
                 random  = ~time_mins,
               data = new_minidf)
   
    mond = lcmm(metab ~time_mins, random  = ~time_mins, mixture = ~time_mins, 
                   ng= as.numeric(class_lat[c]), B = init,
               subject = "study_id", link = as.character(best_link[i]),data = new_minidf)
      
    output_aic_f[[i]][c]  <-  mond$AIC


  }
}

output_aic_f


save(output_aic_f, file="output_aic_f.RData")
load("~/Documents/repos/Dissertation_F/Dissertation-F/output_aic_f.RData")


min_aic_class  = list()
best_df =  data.frame(matrix(NA, ncol = length(output_aic_f)))
colnames(best_df) = paste("M-", 1:length(positions),sep = "")
for(i in 1:length(output_aic_f)){
    names(output_aic_f[[i]]) <- class_lat
    min_aic_class[[i]] = which.min(output_aic_f[[i]])
    best_df[,i] =  paste(best_link[[i]], class_lat[as.numeric(min_aic_class[[i]])],sep = "-")
}
best_df 
class(best_df)
save(best_df, file="labels_summary_FINAL.RData")
load("~/Documents/repos/Dissertation_F/Dissertation-F/labels_summary_FINAL.RData")




subset(best_df, grepl('^beta', 
                      best_df))
best_df[,grepl('^beta', best_df)]

class_two = best_df[,grepl('2$', best_df)]
length(class_two)
beta2 = class_two[,grepl('^beta', class_two)]
length(beta2)

spline2= class_two[,grepl('^splines', class_two)]
length(splines2)



class_three  = best_df[,grepl('3$', best_df)]
length(class_three)
beta3= class_three[,grepl('^beta', class_three)]
length(beta3)
spline3 = class_three[,grepl('^splines', class_three)]
length(spline3)+length(spline2)
length(beta2)+length(beta3)

best_link_f = as.numeric(gsub('[[:alpha:]]+[-]',"", best_df))
length(best_link_f)
length(positions)
best_link[1]

```

#final 
```{r}
output_final = vector(mode="list", length=length(positions))
for(i in 1:length(positions)){
     names(new_minidf)[3] <- "metab"
     new_minidf[,3] = final_df5[i]
     init = lcmm(metab ~time_mins, subject = "study_id", link = as.character(best_link[i]),
                 random  = ~time_mins,
               data = new_minidf)
   
    mond = lcmm(metab ~time_mins, random  = ~time_mins, mixture = ~time_mins, 
                   ng= as.numeric(best_link_f[i]), B = init,
               subject = "study_id", link = as.character(best_link[i]),data = new_minidf)
      
      output_final[[i]]  <-  mond
}

 output_final
# save(output_final, file="output_final.RData")
#load("~/Documents/repos/Dissertation_F/Dissertation-F/output_final.RData")


```


```{r}
#load("~/Documents/repos/Dissertation_F/Dissertation-F/labels_summary_FINAL.RData")
best_df

#load("~/Documents/repos/Dissertation_F/Dissertation-F/output_final.RData")
output_final

new_minidf
```




*Metab 01:metab == meal03._diaminopimelic_acid*
```{r}
#mod 1
init1.0 = lcmm(metab ~time_mins, subject = "study_id", link = "beta",
                 random  = ~time_mins,
               data = new_minidf)
summary(init1.0)

#mod 2
init2.0  = lcmm(metab ~time_mins, subject = "study_id", link = "splines",
                 random  = ~time_mins,
               data = new_minidf)

  

summary(init2.0)


#mond3
init3.0 = lcmm(metab ~time_mins, subject = "study_id", link = "3-quant-splines",
                 random  = ~time_mins,
               data = new_minidf)

mond3 = lcmm( metab~time_mins, random  = ~time_mins, mixture = ~time_mins, 
                   ng= 2, B = init3.0,
               subject = "study_id", link = "beta",data = new_minidf)

summary(mond3)


#mond4
init4.0  = lcmm(metab ~time_mins, subject = "study_id", link = "5-quant-splines",
                 random  = ~time_mins,
               data = new_minidf)





init_linkfunc =summarytable(init1.0, init2.0,init3.0, init4.0 ,which = c("loglik", "conv", "npm", "AIC"))
which.min(init_linkfunc[,4])
order(init_linkfunc[,4])

rep("c",12)
col <- rainbow(10)
#cols = viridis(5)
#library("viridis") 
plot(init1.0, which="linkfunction", bty='l', ylab= "meal03._diaminopimelic_acid", col=col[1], lwd=2, xlab="underlying latent process Trouble") #beta= red
plot(init2.0, which="linkfunction", add=TRUE, col=col[2], lwd=2) #splines - orange
plot(init3.0, which="linkfunction", add=TRUE, col=col[4], lwd=2) #3quant splines green
plot(init4.0, which="linkfunction", add=TRUE, col=col[7], lwd=2)#5 quantsplines - dark blue

legend(x="topleft", legend=c("beta", "splines","splines (3equidistant)","splines (5 at quantiles)"), lty=1, col= c(col[1], col[2],col[4],col[7]), bty="n", lwd=2)
linkbeta <- predictlink(init1.0, ndraws = 2000)
plot(linkbeta, add = TRUE, col = col[4], lty = 2)
legend(legend = c("95% confidence bands", "for splines at quantiles"),
           x = "left", lty = c(2, NA), col = c(col[4], NA), bty = "n", lwd = 1)



#The best link function based on model ini1.0 

#mod 1: 2 latent classes with beta latent process
init1.0 = lcmm(metab ~time_mins, subject = "study_id", link = "beta",
                 random  = ~time_mins,
               data = new_minidf)
summary(init1.0)

two_class_metab1 = lcmm( metab~time_mins, random  = ~time_mins, mixture = ~time_mins, 
                   ng= 2, B = init1.0,
               subject = "study_id", link = "beta",data = new_minidf)

summary(two_class_metab1)
postprob(two_class_metab1)
#mod 2: 3 latent classes with beta latent process 
init2.0  = lcmm(metab ~time_mins, subject = "study_id", link = "beta",
                 random  = ~time_mins,
               data = new_minidf)

    

three_class_metab1 = lcmm( metab~time_mins, random  = ~time_mins, mixture = ~time_mins, 
                   ng= 3, B = init2.0,
               subject = "study_id", link = "beta",data = new_minidf)

summary(three_class_metab1)
postprob(three_class_metab1)


mult_classA = summarytable(two_class_metab1,three_class_metab1,which = c("G","loglik", "npm", "AIC", "%class"))

  #summarytable(two_class_metab1,three_class_metab1,which = c("loglik", "conv", "npm", "AIC"))
which.min(mult_classA[,3])
order(mult_classA[,3])

postprob(two_class_metab1)
postprob(three_class_metab1)
col <- rainbow(10)
#Predictions  based on both best AIC & Link 
mond1pred<- predictlink(two_class_metab1,ndraws=2000)
plot(mond1pred, col=col[4], lty=2, shades=TRUE)
legend(x="left", legend=c("95% confidence bands","for splines at quantiles"),lty=c(2,NA), col=c(col[4],NA), bty="n", lwd=1, cex=0.8)
plot(two_class_metab1,cex.main=0.9)


#Class trajectories based on both best AIC & Link 
new_data_time= data.frame(timepred= seq(30,240,by=5)) ##  USE THIS FROM NOW ON 
new_data_time$time_mins = new_data_time$timepred

predG1 = predictY(init1.0,newdata= new_data_time, var.time = "timepred")
predG2=predictY(two_class_metab1,newdata= new_data_time, var.time = "timepred")
predG3=predictY(three_class_metab1,newdata= new_data_time, var.time = "timepred")

par(mfrow = c(1,3))
plot(predG1, col = 1, lty = 1, lwd = 2, ylab = "Metab 01", legend = NULL, main = "Predicted trajectories G=1", xlim = c(0,300))
plot(predG2, col = c("red", "navy"), lty = 1, lwd = 2, ylab = "Metab 01", legend = NULL,main = "Predicted trajectories G=2", xlim = c(0,300))
plot(predG3, col = 2:4, lty = 1, lwd = 2, ylab = "Metab 01", legend = NULL,main = "Predicted trajectories G=3", xlim = c(0,300))




###OTHER METHOD FOR CLASS SPECIFIC TRAJ with CI bands: 3 LC
predG1.1 = predictY(init1.0,newdata= new_data_time, var.time = "timepred", draws = T)

predG2.2=predictY(two_class_metab1,newdata= new_data_time, var.time = "timepred", draws = T)
predG3.3=predictY(three_class_metab1,newdata= new_data_time, var.time = "timepred", draws = T)


# plot(predG2,lwd = c(3, 1),xlab = "Time Minutes: Simple Metab 01", ylab = "Latent Process: Simple Metab 01")

par(mfrow = c(1,1))
plot(predG2.2, col = c("red", "navy"), lty = 1, lwd = 2, ylab = "Metab 01", legend = NULL)
legend(x = "bottomright", legend = c("class 1", "Time Lower ", "time upper", "class 2","Time Lower ", "time upper" ), col = c(rep("red",3), rep("navy",3)), lwd = 2, lty = c(0,1,2,0,1,2), ncol =2, bty = "n", cex = 0.7)


plot(predG3.3, lwd = c(3, 1),xlab = "Time Minutes: Simple Metab 01", ylab = "Metab 01", c("red", "navy", "green"), legend = NULL)
legend(x = "bottomright", legend = c("class 1", "Time Lower ", "time upper", "class 2","Time Lower ", "time upper", "class 3","Time Lower ", "time upper"  ), col = c(rep("red",3), rep("navy",3),rep("green",3)), lwd = 2, lty = c(0,1,2,0,1,2,0,1,2), ncol =3, bty = "n", cex = 0.7)




par(mfrow = c(1,3))
plot(predG1.1, col = 1, lty = 1, lwd = 2, ylab = "Metab 01", legend = NULL, main = "Predicted trajectories G=1", xlim = c(0,300))
plot(predG2.2, col = c("red", "navy"), lty = 1, lwd = 2, ylab = "Metab 01", legend = NULL,main = "mean  Predicted trajectories G=2", xlim = c(0,300))
plot(predG3.3, col = 2:4, lty = 1, lwd = 2, ylab = "Metab 01", legend = NULL,main = "meanPredicted trajectories G=3", xlim = c(0,300))


postprob(two_class_metab1)

#Weighted mean of predicted trajectories based on both best AIC & Link 
par(mfrow = c(1,1))
plot(three_class_metab1, which = "fit", var.time = "time_mins", xlab = "Time in Minutes: Simple Metab 01",
     bty = "l", break.times = 5, ylab = "three latent process", lwd = 2, marg = FALSE,
     ylim = c(-8, 2))


par(mfrow = c(1,1))
plot(two_class_metab1, which = "fit", var.time = "time_mins", xlab = "Time in Minutes",
     bty = "l", break.times = 6, ylab = "two latent process", lwd = 2, marg = FALSE,
     ylim = c(-8, 2))




#Choosing the best model: Label switching look 

label_switching = summarytable(init1.0,two_class_metab1,three_class_metab1, which = c("G", "loglik", "conv", "npm", "AIC", "BIC", "SABIC", "entropy","ICL", "%class"))



summaryplot(init1.0,two_class_metab1,three_class_metab1,which = c("AIC", "BIC", "entropy","ICL"))

order(label_switching[,5])
order(label_switching[,7])








```




*Metab 02:meal03._alpha_l_fucose*
```{r}

#mod 1
init1.0b = lcmm(meal03._alpha_l_fucose ~time_mins, subject = "study_id", link = "beta",
                 random  = ~time_mins,
               data = new_minidf)
summary(init1.0b)
#mod 2
init2.0b  = lcmm( meal03._alpha_l_fucose ~time_mins, subject = "study_id", link = "splines",
                 random  = ~time_mins,
               data = new_minidf)

    
#mond3
init3.0b = lcmm( meal03._alpha_l_fucose ~time_mins, subject = "study_id", link = "3-quant-splines",
                 random  = ~time_mins,
               data = new_minidf)


#mond4
init4.0b  = lcmm(meal03._alpha_l_fucose  ~time_mins, subject = "study_id", link = "5-quant-splines",
                 random  = ~time_mins,
               data = new_minidf)



init_linkfunc_b =summarytable(init1.0b, init2.0b,init3.0b, init4.0b ,which = c("loglik", "conv", "npm", "AIC"))
which.min(init_linkfunc_b[,4])
order(init_linkfunc_b[,4])


col <- rainbow(10)

plot(init1.0b, which="linkfunction", bty='l', ylab= "meal03._alpha_l_fucose", col=col[1], lwd=2, xlab="underlying latent process Trouble") #beta= red
plot(init2.0b, which="linkfunction", add=TRUE, col=col[2], lwd=2) #splines - orange
plot(init3.0b, which="linkfunction", add=TRUE, col=col[4], lwd=2) #3quant splines green
plot(init4.0b, which="linkfunction", add=TRUE, col=col[7], lwd=2)#5 quantsplines - dark blue

legend(x="topleft", legend=c("beta", "splines","splines (3equidistant)","splines (5 at quantiles)"), lty=1, col= c(col[1], col[2],col[4],col[7]), bty="n", lwd=2)


#The best link function based on model ini1.0 

#mod 1
init2.0b  = lcmm( meal03._alpha_l_fucose ~time_mins, subject = "study_id", link = "splines",
                 random  = ~time_mins,
               data = new_minidf)

summary(init2.0b)

two_class_metab2 = lcmm( meal03._alpha_l_fucose ~time_mins, random  = ~time_mins, mixture = ~time_mins, 
                   ng= 2, B = init2.0b,
               subject = "study_id", link = "splines",data = new_minidf)

summary(two_class_metab2)
postprob(two_class_metab2)
#mod 2
#mod 1
init2.0b  = lcmm( meal03._alpha_l_fucose ~time_mins, subject = "study_id", link = "splines",
                 random  = ~time_mins,
               data = new_minidf)

summary(init2.0b)
    

three_class_metab2 = lcmm( metab~time_mins, random  = ~time_mins, mixture = ~time_mins, 
                   ng= 3, B = init2.0b,
               subject = "study_id", link = "splines",data = new_minidf)

summary(three_class_metab2)
postprob(three_class_metab2)

mult_classB =summarytable(two_class_metab2,three_class_metab2,which = c("G","loglik", "npm", "AIC", "%class"))
which.min(mult_classB[,3])
order(mult_classB[,4])

col <- rainbow(10)
#Predictions  based on both best AIC & Link 
mond1pred_b<- predictlink(two_class_metab2,ndraws=2000)
plot(mond1pred_b, col=col[4], lty=2, shades=TRUE)
legend(x="left", legend=c("95% confidence bands","for splines at quantiles"),lty=c(2,NA), col=c(col[4],NA), bty="n", lwd=1, cex=0.8)
plot(two_class_metab2,cex.main=0.9)


#Class trajectories based on both best AIC & Link 
new_data_time= data.frame(timepred= seq(30,240,by=5)) ##  USE THIS FROM NOW ON 
new_data_time$time_mins = new_data_time$timepred
pred_meta_1 = predictY(init2.0b,newdata= new_data_time, var.time = "timepred")
pred_met_2=predictY(two_class_metab2,newdata= new_data_time, var.time = "timepred")
pred_met_3=predictY(three_class_metab2,newdata= new_data_time, var.time = "timepred")






par(mfrow = c(1,3))
plot(pred_meta_1, col = 1, lty = 1, lwd = 2, ylab = "Metab 01", legend = NULL, main = "Predicted trajectories G=1", xlim = c(0,300))
plot(pred_met_2, col = c("red", "navy"), lty = 1, lwd = 2, ylab = "Metab 01", legend = NULL,main = "mean  Predicted trajectories G=2", xlim = c(0,300))
plot(pred_met_3, col = 2:4, lty = 1, lwd = 2, ylab = "Metab 01", legend = NULL,main = "meanPredicted trajectories G=3", xlim = c(0,300))



#Weighted mean of predicted trajectories based on both best AIC & Link 
class(predG2$times)

plot(pred_met_2,lwd = c(3, 1),xlab = "Time Minutes: Trouble Metab 02", ylab = "Latent Process: Trouble  Metab 02")
plot(pred_met_3, lwd = c(3, 1),xlab = "Time Minutes: Trouble Metab 02", ylab = "Latent Process: Trouble Metab 02")



#Goodness of Fit || col= c(col[2], col[4], col[7])
plot(three_class_metab2, which = "fit", var.time = "time_mins", xlab = "Time in Minutes: Complex Metab 02",
     bty = "l", break.times = 5, ylab = "latent process", lwd = 2, marg = FALSE,
     ylim = c(-5, 2))


label_switching2 = summarytable(init2.0b,two_class_metab2,three_class_metab2, which = c("G", "loglik", "conv", "npm", "AIC", "BIC", "SABIC", "entropy","ICL", "%class"))

summaryplot(init2.0b,two_class_metab2,three_class_metab2,which = c("AIC", "BIC", "entropy","ICL"))
label_switching2
order(label_switching2[,5])
order(label_switching2[,6])
order(label_switching2[,7])






```





















